# Practical Machine Learning Course Assignment

## Executive Summary

Using data collected from various fitness tracking devices (Fitbit, Jawbone, etc) the goal of this project is to use this data, which has been provided as part of the assignment, to predict the manner in which the participants performed a particular exercise. Data from accelerometers on the belt, forearm and dumbbell of six participants will be used to predict this dependent value denoted as "classe"

## Data Loading and Cleaning

The data can be found at the following locations:
<br/>

* Training
<br/>
<a href="https://d396qusza40orc.cloudfront.net/predmachlearn/pml-training.csv">https://d396qusza40orc.cloudfront.net/predmachlearn/pml-training.cs</a>
<br/>

* Test
<br/>
<a href="https://d396qusza40orc.cloudfront.net/predmachlearn/pml-testing.csv">https://d396qusza40orc.cloudfront.net/predmachlearn/pml-testing.csv</a>
<br/><br/>

```{r echo=TRUE}
## Load the data
set.seed(1234)
train <- read.csv("pml-training.csv", na.strings=c("","NA","NULL"), strip.white = T)
test <- read.csv("pml-testing.csv", na.strings=c("","NA","NULL"), strip.white = T)
dim(train);dim(test)
```

<br/>Get rid of rows with NAs
<br/>

```{r echo=TRUE}
## Clean the data
naRows <- train[, apply(train, 2, function(x) { !any(is.na(x))})]
naRows2 <- test[, apply(test, 2, function(x) { !any(is.na(x))})]
dim(naRows); dim(naRows2)
```

<br/>Get rid of non-meaningul columns
<br/>

```{r echo=TRUE}
train <- naRows[, -c(1:8)]
test  <- naRows2[, -c(1:8)]
dim(train);dim(test)
```

<br/>Now lets use the method taught in class to partition the training data into training and test sets
<br/>

```{r echo=TRUE}
library(caret)
library(randomForest)
inTrain <- createDataPartition(train$classe, p = 0.7, list = F)
training <- train[inTrain,]
validating <- train[-inTrain,]
dim(training); dim(validating)
```

## Random Forest

```{r echo=TRUE, cache=TRUE}
## Takes a looong time, so caching is on
rf <- train(classe ~ . , data = training, model = "rf", importance = T)
## Predict against the testing data
pd <- predict(rf, newdata = validating)
```

## Cross Validation
<br/>
Check how well the rf model predicts against the vaidation data

```{r echo=TRUE}
sum(pd == validating$classe) / length(pd)
```

```{r echo=TRUE}
confusionMatrix(validating$classe, pd)$table
```

99.28% accuracy and the confusion matrix looks good...
<br/>
<br/>
Produce the values used to submit for assignment quiz

```{r echo=TRUE}
predict(rf, newdata=test)
```

Let's plot the variable importance (top 20)

```{r echo=TRUE}
varImpPlot(rf$finalModel, n.var = 20)
```
